---
title: "portfolio"
output: html_document
date: "2023-04-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
df <- read_excel("/Users/lindsaystall/Downloads/PIRUS_Public_May2020.xlsx")
df[df == -99 | df == -88] <- NA
```


```{r}
df2 <- data.frame(table(df$Previous_Criminal_Activity))

ggplot(df2, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "pink", color = "black") +
  xlab("Previous Criminal Activity") +
  ylab("Count") +
  ggtitle("Previous Criminal Activity for Females") +
  theme_bw()
```

0 = No previous criminal activity
1 = Previous (non-violent) minor criminal activity (e.g., convicted of a misdemeanor crime) 
2 = Previous (non-violent) serious criminal activity (e.g., convicted of a felony crime)
3 = Previous violent crime

Most terrorists in this data set have no previous criminal activity

```{r}
df %>% 
  filter(!is.na(Previous_Criminal_Activity_Type1)) %>% 
  group_by(Previous_Criminal_Activity_Type1) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))
```

1 = Homicide
2 = Forcible Rape
3 = Robbery
4 = Aggravated Assault 
5 = Burglary
6 = Larceny-Theft
7 = Motor Vehicle Theft 
8 = Arson
9 = Simple Assault
10 = Fraud
11 = Forgery
12 = Embezzlement
13 = Driving Under the Influence (DUI)/Driving While Intoxicated (DWI)
14 = Prostitution
15 = Vandalism
16 = Drug related (e.g., selling/distributing drugs, cultivating drugs, but not simply drug use).
17 = Parole violation
18 = Unlawful possession, transportation, or use of a firearm 
19 = Domestic violence/spousal abuse
20 = Other

Of the terrorists with previous charges, the most common charges was Aggravated Assault, Robbery, and Simple Assault. 


```{r}
psych_df <- data.frame(count = c(
  sum(df$Psychological == 0, na.rm = TRUE),
  sum(df$Psychological == 1, na.rm = TRUE),
  sum(df$Psychological == 2, na.rm = TRUE)
), 
value = c("No", "Speculation", "Professionally Diagnosed"))

ggplot(data = psych_df, aes(x="", y=count, fill=value)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  labs(fill="Psychological") +
  theme_void()
```

The vast majority of terrorists do not have a reported or speculated mental illness. I know this is very much in line with existing literature.


```{r}
counts <- table(df$Social_Stratum_Childhood)

ggplot(data.frame(Social_Stratum_Childhood = names(counts), counts), aes(x = Social_Stratum_Childhood, y = counts)) +
  geom_bar(stat = "identity", fill = "pink", color = "black") +
  ggtitle("Social Stratum in Childhood") +
  xlab("Social Stratum") +
  ylab("Count")
```

1 = Low (e.g. receives welfare, lives close to the poverty line, regularly unemployed or at best works a blue collar job, lives in subsidized housing)
2 = Middle (e.g. does not receive welfare, lives in lower-middle or middle class neighborhood, has steady professional employment, owns or holds a mortgage on a house, has college degree)
3 = High (e.g. works a high-income, white-collar job, lives and owns a house in a middle or upper class neighborhood, can afford luxury items, has college degree or is self-employed as a successful entrepreneur)


Most terrorists grew up middle class.


```{r}
counts2 <- table(df$Social_Stratum_Adulthood)

ggplot(data.frame(Social_Stratum_Adulthood = names(counts2), counts2), aes(x = Social_Stratum_Adulthood, y = counts)) +
  geom_bar(stat = "identity", fill = "pink", color = "black") +
  ggtitle("Social Stratum in Adulthood") +
  xlab("Social Stratum") +
  ylab("Count")
```

Most terrorists remained middle class into adulthood

```{r}
table(df$Alcohol_Drug)
```

The vast majority of terrorists (89%) do not have a alcohol or drug problem.


```{r}
ggplot(df, aes(x = Absent_Parent)) +
  geom_histogram(binwidth = 1, fill = "pink", color = "black") +
  labs(x = "Absent Parent", y = "Count", title = "Histogram of Absent Parent")
```
0 = No
1 = Yes, absent mother 
2 = Yes, absent father 
3 = Yes, both absent

Most terrorists did not have absent parents.

```{r}
ggplot(df, aes(x = Close_Family)) +
  geom_histogram(binwidth = 1, fill = "pink", color = "black") +
  labs(x = "Family closeness", y = "Count", title = "Histogram of AFamily Closeness")
```

1 = Distant (has little regular contact with most family members, rarely attends family events, does not celebrate holidays with family, etc.)
2 = Close (regularly interacts with one or more family members, attends family events on a regular basis, usually celebrates holidays with family, etc.)


Most terrorists have close relationships with family!!!

